---
title: "data cleaning of netflix_titles.csv for netflixListings"
output: html_notebook
---

purpose of notebook:
  (/) initial look at data to get a basic understanding, gather todos
  (/) tidy up and cleaning of data set
  (x) save pre-processed data set -> not possible as long as not flat, ie, tibble contains lists as entries

todos:
  (/) check NAs with rtatman visualization 
  (/) check for duplicated rows with rtatman viz
  (/) column names look fine, no need for clean_names() from janitor
  (/) names in column cast, description need other codec -> use read_csv is fine, but read.csv() seems not to use UTF-8
  (/) missing values in director -> filled with NA by read_csv, likely not missing at random, could be checked with additional database like IMDB
  (/) dates instead of chr in column date
  (/) cast column is a very long string, should be a list of names or different encoding as needed
  (/) in duration remove the unit, may need to be separated for TV Show/Seasons and Movie/min later
  (/) listed_in should be a list of categories, better different encoding as needed, also may need cleaner naming convention
  (/) column country has some times multiple countries, replace str with list
  (?) change categorical encoding for: type, country, rating, listed_in -> may be later when required
  
observations:
  (i) columns: show_id -> chr, unique values, identifier - Unique ID for every Movie / Tv Show
               type -> chr, categorical, 2 values - Identifier - A Movie or TV Show
               title -> chr, (almost?) unique value - Title of the Movie / Tv Show
               director -> chr, categorical, many values - Director of the Movie
               cast -> chr, list of many names - Actors involved in the movie / show
               country -> chr, categorical - Country where the movie / show was produced
               date_added -> date, time value - Date it was added on Netflix
               release_year -> int, categorical - Actual Release year of the move / show
               rating -> chr, categorical - TV Rating of the movie / show 
               duration -> int with unit, two main features (TV Show in Seasons, Movie in min) - Total Duration - in minutes or number of seasons
               listed_in -> chr, categorical, list of categories - Genere
               description -> chr, longer text - The summary description
  (i) size of data set is 7787 rows
  (i) no duplicated rows
  (i) missing values: a lot in directors, some in cast and country -> only imputation possible is by using another data set like IMDB
  
stories:
  (!) description with NLP sentiment analysis
  (!) network analysis with cast and director
  (!) increase of number over all, types, countries, rating, listed_in; usage of cast or director,  .. over time 
               

---
load packages
```{r}
library(tidyverse) # tidy data frame
library(lubridate) # date 
library(stringr) # strings, but also already in tidyverse
library(scrubr) # like dplyr but specifically for occurrence data
```

---
import data to tibble named 'netflix'
```{r}
netflix_raw <- read_csv(file = '../data/netflix_titles.csv')
netflix <- netflix_raw
```

---
check for missing values in data set 
```{r, fig.asp=1.5}
# create a data frame with information on whether the value in each cell is missing
missing_by_column <- netflix %>%
  is.na %>% # check if each cell is na
  as_tibble %>% # convert to data frame
  mutate(row_number = 1:nrow(.)) %>% # add a column with the row number
  gather(variable, is_missing, -row_number) # turn wide data into narrow data

# plot the missing values in our data frame, with a good-looking theme
ggplot(missing_by_column, aes(x = variable, y = row_number, fill = is_missing)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_grey(name = "",
                  labels = c("present", "missing")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.9, size=8)) +
  labs(x = "variables in dataset",
       y = "rows / oberservations")
```

---
check for duplicated rows
```{r}
# get the row numbers of duplictaed rows
duplicated_rows <- tibble(duplicated = duplicated(netflix), row = 1:nrow(netflix)) %>%
  filter(duplicated == T)

# plot duplicated rows as black lines
ggplot(duplicated_rows, aes(xintercept = row)) +
  geom_vline(aes(xintercept = row)) + # plot a black line for each duplicated row
  ggtitle("indexes of duplicated rows") + # add title
  coord_flip() + scale_x_reverse() # flip x & y axis and reverse the x axis
```

---
data cleaning section
---

dates instead of chr in column date
```{r}
dates <- netflix %>%
  select(date_added) %>%
  mutate(date_added_format = mdy(date_added))

netflix$date_added <- dates$date_added_format

print(netflix$date_added, max=7)
```


cast column is a very long string, should be a list of names
```{r}
cast <- netflix %>%
  select(cast) %>%
  mutate(cast_list = str_split(cast, ", "))

netflix$cast <- cast$cast_list

print(netflix$cast, max=2)
print(netflix$cast[1])
```

in duration remove the unit, may need to be separated for TV Show/Seasons and Movie/min later
```{r}
duration <- netflix %>%
  select(duration) %>%
  mutate(duration_unit = parse_number(duration))

netflix$duration <- duration$duration_unit
```

listed_in should be a list of categories, also needs cleaner naming convention
```{r}
remove_comma <- function(str) gsub(",", "", gsub("([a-zA-Z]),", "\\1", str)) # function to remove commas from a given str
glue_ampersand <- function(str) gsub(" & ", "_", str) # better use _ instead of &
remove_apos <- function(str) gsub("'", "", gsub("([a-zA-Z])'", "\\1", str)) # function to remove apostrophes from a given str

listed_clean <- netflix %>%
  select(listed_in) %>%
  mutate(listed_clean = remove_comma(listed_in)) %>% # remove commas from str
  mutate(listed_clean = str_remove_all(listed_clean, "International|TV|Shows|Movies|Series|Features")) %>% # remove stop words from listed_in column
  mutate(listed_clean = str_trim(listed_clean)) %>% # remove whitespace from start and end
  mutate(listed_clean = str_squish(listed_clean)) %>% # remove repeated whitespace inside str
  mutate(listed_clean = glue_ampersand(listed_clean)) %>% # remove whitspace around ampersand 
  mutate(listed_clean = remove_apos(listed_clean)) %>% # remove apostrophes from str
  mutate(listed_clean = str_split(listed_clean, " ")) # convert to list

netflix$listed_in <- listed_clean$listed_clean

print(netflix$listed_in, max=2)
```
column country has some times multiple countries, replace str with list
```{r}
remove_last_comma <- function(str) gsub(",$", "", str) # function to remove last comma from a given str

country <- netflix %>%
  select(country) %>%
  mutate(country_list = remove_last_comma(country)) %>%
  mutate(country_list = str_split(country_list, ", "))
  
netflix$country <- country$country_list

print(netflix$country, max=2)
```

column director has some times multiple names, replace str with list
```{r}
director <- netflix %>%
  select(director) %>%
  mutate(director_list = str_split(director, ", "))

netflix$director <- director$director_list

print(netflix$director, max=2)
```

--- 
write cleaned netflix data to csv -> not possible since not flat
```{r}
# write_csv(netflix, '../data/netflix_clean.csv')
```

---
overview of data with head, tail, summary
```{r}
head(netflix, 10)
```

```{r}
tail(netflix, 10)
```

```{r}
summary(netflix)
```

