---
title: "describe and visualize of US weather event data"
subtitle: "city selection - multivariate"
output: html_notebook
---

---
purpose of notebook
---

  (-) describe & visualize relationship between variables (multivariate)
  (-) correlation plots 
  (-) gather interesting observations for further investigation
  (-) gather possible new features for extraction
  
todos:
  (-) ...

---
information
---
https://www.kaggle.com/sobhanmoosavi/us-weather-events
https://smoosavi.org/datasets/lstw

Moosavi, Sobhan, Mohammad Hossein Samavatian, Arnab Nandi, Srinivasan Parthasarathy, and Rajiv Ramnath. “Short and Long-term Pattern Discovery Over Large-Scale Geo-Spatiotemporal Data.” In Proceedings of the 25th ACM SIGKDD International Conference on Knowledge Discovery & Data Mining, ACM, 2019.

Weather event is a spatiotemporal entity, where such an entity is associated with location and time. Following is the description of available weather event types:

    Severe-Cold: The case of having extremely low temperature, with temperature below -23.7 degrees of Celsius.
    Fog: The case where there is low visibility condition as a result of fog or haze.
    Hail: The case of having solid precipitation including ice pellets and hail.
    Rain: The case of having rain, ranging from light to heavy.
    Snow: The case of having snow, ranging from light to heavy.
    Storm: The extremely windy condition, where the wind speed is at least 60 km/h.
    Other Precipitation: Any other type of precipitation which cannot be assigned to previously described event types.

The weather data is provided in terms of a CSV file with the following attributes:
      Attribute	        Description	                                                Nullable
1	    EventId	          This is the identifier of a record	                        No
2   	Type	            The type of an event; examples are rain and snow.	          No
3	    Severity	        The severity of an event, wherever applicable.        	    Yes
4	    StartTime (UTC)	  The start time of an event in UTC time zone.          	    No
5	    EndTime (UTC)	    The end time of an event in UTC time zone.	                No
6	    TimeZone	        The US-based timezone based on the location of an event     No
                        (eastern, central, mountain, and pacific).	
7	    LocationLat	      The latitude in GPS coordinate.	                            Yes
8   	LocationLng	      The longitude in GPS coordinate.	                          Yes
9	    AirportCode	      The airport station that a weather event is reported from.	Yes
10	  City	            The city in address record.	                                Yes
11	  County	          The county in address record.	                              Yes
12	  State	            The state in address record.	                              Yes
13	  ZipCode	          The zipcode in address record.	                            Yes

---
insights 
---
  
  (i) 
  
---
stories
---

  (!) compare 4 big cities which are all close to the coast but vary in significantly in longitude and latitude

---
load packages
---
```{r load packages, include=FALSE}
library(tidyverse) # tidy data frame
library(ggthemes) # for extra plot themes
library(plotly) # make ggplots interactive
library(lubridate) # functions to work with date-times and time-spans
library(corrplot) # correlation plots
```

---
overview
---
```{r}
head(weather)
```
```{r}
summary(weather)
```

---
multivariate City
---





