---
title: "cleaning for plastic waste makers index data"
output: html_notebook
---

---
purpose of notebook
---

  (/) initial look at data to get a basic understanding and gather todos
  (/) tidy up and cleaning of data set
  (x) save pre-processed data set

todos:
  (-) ...

---
information
---

name: makeovermonday_2021w22
link: https://data.world/makeovermonday/2021w22
title: 2021/W22: The Plastic Waste Makers Index
Data Source: [Minderoo](https://www.minderoo.org/plastic-waste-makers-index/data/indices/producers/) from 2019

---
observations
---
  
  (i) columns: rank                                         numeric, ordered, unique, can serve as identifier, rank of producer according to index
               polymer_producer                             string, unique identifier, name of producer
               no_of_assets                                 numeric, metric, number of assets of the producer
               production_of_in_scope_polymers              numeric, metric in million metric tons, production of plolymers that are in-scope of preceding analysis
               flexible_format_contribution_to_sup_waste    numeric, metric in million metric tons, flexible form of contribution to sup waste
               rigid_format_contribution_to_sup_waste       numeric, metric in million metric tons, rigid form of contribution to sup waste
               total_contribution_to_sup_waste              numeric, metric in million metric tons, total contribution is the sum of flexible and rigid
  (i) no missing values at all, also it is a very small dataset
  (i) no duplicated rows
  (i) no changes were made to data set
  (i) added a variable which gives the percentage of the sup_waste from the produced polymers
               
---
load packages
---
```{r load packages, include=FALSE}
library(tidyverse) # tidy data frame
library(lubridate) # functions to work with date-times and time-spans
library(scrubr) # like dplyr but specifically for occurrence data
```

---
import data
---
```{r}
plastic <- read_csv(file = '../data/plastic_waste_makers_index.csv')
```

---
first look at data
---
```{r}
head(plastic, 25)
```
```{r}
tail(plastic)
```
```{r}
summary(plastic)
```

---
missing values
---
no missing values at all, also it is a very small dataset

```{r}
#create data frame with information on whether the value in each cell is missing
missing_by_column <- plastic %>%
  is.na %>% # check is each cell is na
  as_tibble %>% # convert to data frame
  mutate(row_number = 1:nrow(.)) %>% # add a column with the row number
  gather(variable, is_missing, -row_number) # turn wide data into narrow data

# plot the missing values in our data frame
ggplot(missing_by_column, aes(x = variable, y = row_number, fill =  is_missing)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_grey(name = "",
                  labels = c("present", "missing")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=8)) +
  labs(x = "vairables in dataset",
       y = "rows / observations")
```

---
duplicated rows
---
no duplicated rows

```{r}
# get row number of duplicated rows
duplicated_rows = tibble(duplicated = duplicated(plastic), row = 1:nrow(plastic)) %>%
  filter(duplicated == T)

# plot duplicated rows as black lines
ggplot(duplicated_rows, aes(xintercept = row)) +
  geom_vline(aes(xintercept = row)) + # plot a black line for each duplicated row
  ggtitle("indexes of duplicated rows") + # add title
  coord_flip() + # flip x and y axis
  scale_x_reverse() # reverse x axis
```

---
cleaning
---

nothing to clean 

---
additional variables
---
added a variable which gives the percentage of the sup_waste from the produced polymers

```{r}
plastic <- plastic %>%
  mutate(total_waste_div_production = round(total_contribution_to_sup_waste / production_of_in_scope_polymers ,2))
```

---
save processed data
---

no changes were made to data set


















