---
title: "describe univariate for customer personality analysis data"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/uni_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] describe & visualize single variables (univariate)
  - [ ] gather interesting observations for further investigation
  - [ ] gather possible new features for extraction
  
  
## insights 

  - male_circumcision_percent: it seems that there are two populations of countries, one of mostly circumized males, the other of mostly not-circumzied males; but also the few countries in the middle around 50% are interesting 
  

## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(plotly) # make ggplots interactive
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/marketing_campaign_clean.csv')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## univariate numeric ID

- ID: 

```{r}
# one variable, continuous x, show distribution
name = 'ID'
tmp_df <- df %>% rename(value = ID) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```



## univariate numeric Year_Birth

- Year_Birth: bimodal? compare to baseline of overall population? 

```{r}
# one variable, continuous x, show distribution
name = 'Year_Birth'
tmp_df <- df %>% rename(value = Year_Birth) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical Education

- Education: 

```{r}
# one variable, categorical x, show distribution
name = 'Education'
tmp_df <- df %>% rename(value = name) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical Marital_Status

- 

```{r}
# one variable, categorical x, show distribution
name = 'Marital_Status'
tmp_df <- df %>% rename(value = name) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric Income

- 

```{r}
# one variable, continuous x, show distribution
name = 'Income'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical Kidhome

- 

```{r}
# one variable, continuous x, show distribution
name = 'Kidhome'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```

```{r}
# one variable, categorical x, show distribution
name = 'Kidhome'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical Teenhome

-

```{r}
# one variable, categorical x, show distribution
name = 'Teenhome'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric Dt_Customer

- 

```{r}
# one variable, continuous x, show distribution
name = 'Dt_Customer'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric Recency

- 

```{r}
# one variable, continuous x, show distribution
name = 'Recency'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntWines

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntWines'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntFruits

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntFruits'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntMeatProducts

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntMeatProducts'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntFishProducts

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntFishProducts'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntSweetProducts

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntSweetProducts'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntGoldProds

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntGoldProds'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical NumDealsPurchases

- 

```{r}
# one variable, categorical x, show distribution
name = 'NumDealsPurchases'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical NumWebPurchases

- 

```{r}
# one variable, categorical x, show distribution
name = 'NumWebPurchases'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical NumCatalogPurchases

- 

```{r}
# one variable, categorical x, show distribution
name = 'NumCatalogPurchases'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical NumStorePurchases

- 

```{r}
# one variable, categorical x, show distribution
name = 'NumStorePurchases'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical NumWebVisitsMonth

- 

```{r}
# one variable, categorical x, show distribution
name = 'NumWebVisitsMonth'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical AcceptedCmp3

- 

```{r}
# one variable, categorical x, show distribution
name = 'AcceptedCmp3'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical AcceptedCmp4

- 

```{r}
# one variable, categorical x, show distribution
name = 'AcceptedCmp4'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical AcceptedCmp5

- 

```{r}
# one variable, categorical x, show distribution
name = 'AcceptedCmp5'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical AcceptedCmp1

- 

```{r}
# one variable, categorical x, show distribution
name = 'AcceptedCmp1'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical AcceptedCmp2

- 

```{r}
# one variable, categorical x, show distribution
name = 'AcceptedCmp2'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical Complain

- 

```{r}
# one variable, categorical x, show distribution
name = 'Complain'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical Response

- 

```{r}
# one variable, categorical x, show distribution
name = 'Response'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric AcceptedCmpAll

- 

```{r}
# one variable, categorical x, show distribution
name = 'AcceptedCmpAll'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% mutate(value = factor(value))

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric MntAll

- 

```{r}
# one variable, continuous x, show distribution
name = 'MntAll'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric PurchasesAll

- 

```{r}
# one variable, continuous x, show distribution
name = 'PurchasesAll'
tmp_df <- df %>% rename(value = name) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```










