---
title: "cleaning for customer personality analysis data"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/clean_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] initial look at data to get a basic understanding and gather todos -> df_raw
  - [ ] tidy up and cleaning of data set -> df
  - [ ] all processing of the data set happens here and is saved to pre-/processed csv-> df
 

## observations

  - no MAR
  - no duplicates
  - cleaning: 
    + categorical variables as factors (even though it can't be saved in csv)
    + correct a typo in basis factor
  - added variables: non yet
               

## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(lubridate) # functions to work with date-times and time-spans
library(janitor) # expedite the initial data exploration and cleaning that comes with any new data set
library(plotly)

library(readxl) # read_excel to read xslx dataset
```


## import data

```{r}
df_raw <- read_tsv('../data/marketing_campaign.csv') 
df <- df_raw
```


## first look at data

```{r}
head(df_raw, 25)
```
```{r}
tail(df_raw)
```
```{r}
summary(df_raw)
```


## missing values

- only some NAs in column income, might be intentional 

```{r fig.height=20, fig.width=10}
#create data frame with information on whether the value in each cell is zero
tmp_df <- df_raw == 0
missing_by_column <- tmp_df %>%
  as_tibble %>% # convert to data frame
  mutate(row_number = 1:nrow(.)) %>% # add a column with the row number
  gather(variable, is_missing, -row_number) # turn wide data into narrow data

# plot the missing values in our data frame
ggplot(missing_by_column, aes(x = variable, y = row_number, fill =  is_missing)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_grey(name = "",
                  labels = c("present", "is zero", "is NA")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10)) +
  labs(x = "vairables in dataset",
       y = "rows / observations")
```


## duplicated rows

- no duplicates

```{r}
# get row number of duplicated rows
duplicated_rows <- tibble(duplicated = duplicated(df_raw), row = 1:nrow(df_raw)) %>%
  filter(duplicated == T)

# plot duplicated rows as black lines
ggplot(duplicated_rows, aes(xintercept = row)) +
  geom_vline(aes(xintercept = row)) + # plot a black line for each duplicated row
  ggtitle("indexes of duplicated rows") + # add title
  coord_flip() + # flip x and y axis
  scale_x_reverse() # reverse x axis
```


## cleaning

- 

```{r}
df <- df %>%
  mutate(Education = factor(Education)) %>% # categorical col as factor
  mutate(Marital_Status = factor(Marital_Status)) %>% # categorical col as factor
  mutate(Dt_Customer = as.Date(Dt_Customer, "%d-%m-%Y")) %>%  # convert chr to date 
  select(-Z_CostContact, -Z_Revenue) # these columns are constant, so we drop them
```


## remove remove anomalies

- 

```{r}
df <- df %>% 
  filter(!((Year_Birth == 1893)|
           (Year_Birth == 1899)|
           (Year_Birth == 1900))) %>% 
  filter(!((Marital_Status == 'Alone')|
           (Marital_Status == 'YOLO')|
           (Marital_Status == 'Absurd'))) %>% 
  filter(!(Income > 120000)) %>% 
  filter(!(MntMeatProducts > 1000)) %>% 
  filter(!(MntSweetProducts > 200)) %>% 
  filter(!(MntGoldProds > 250))
```


## additional variables

- non yet

```{r}
# new variable: sum of all AcceptedCmp*
df <- df %>% 
  rowwise %>% 
  mutate(AcceptedCmpAll = sum(c_across(contains("AcceptedCmp")))) 

# new variable: Mnt*
df <- df %>% 
  rowwise %>% 
  mutate(MntAll = sum(c_across(contains("Mnt")))) 

# new variable: *Purchases
df <- df %>% 
  rowwise %>% 
  mutate(PurchasesAll = sum(c_across(contains("Purchases")))) 
```


```{r}
# check results from cleaning and added variables
head(df)
```
```{r}
summary(df)
```


## save processed data

```{r}
df %>% write_csv(file = '../data/marketing_campaign_clean.csv')
```

