---
title: "cleaning for camera data set"
output: 
  html_document:
    keep_md: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/clean_"
)
```

---
purpose of notebook
---

  (-) initial look at data to get a basic understanding and gather todos
  (-) tidy up and cleaning of data set
  (-) save pre-processed data set

---
observations
---
  
  (i) columns:   
  | # | attribute | description / properties |                                                                                               
  |:-:|:---------:|:-----------------------------------------------------------:|
  | 1 | rank | numeric, ordered, unique, can serve as identifier, rank of producer according to index |
| 2 |       Type      | The type of an event; examples are accident and congestion. |
| 3 |     Severity    | The  severity is a value between 0 and 4, where 0 indicates the least impact  on traffic (i.e., short delay as a result of the event) and 4 indicates  a significant impact on traffic (i.e., long delay). |
| 4 |       TMC       | Each traffic event has a Traffic Message Channel (TMC) code which provides a more detailed description on type of the event.                                                                               |
| 5 |   Description   | The natural language description of an event.                                                                                                                                                              |
| 6 | StartTime (UTC) | The start time of an event in UTC time zone.                                                                                                                                                               |
| 7 |  EndTime (UTC)  | The end time of an event in UTC time zone.                                                                                                                                                                 |
| 8 |     TimeZone    | The US-based timezone based on the location of an event (eastern, central, mountain, and pacific).                                                                                                         |
| 9 |   LocationLat   | The latitude in GPS coordinate.                                                                                                                                                                            |
| 10 |   LocationLng   | The longitude in GPS coordinate.                                                                                                                                                                           |
| 11 |  Distance (mi)  | The length of the road extent affected by the event.                                                                                                                                                       |
| 12 |   AirportCode   | The closest airport station to the location of a traffic event.                                                                                                                                            |
| 13 |      Number     | The street number in address record.                                                                                                                                                                       |

               rank                                         numeric, ordered, unique, can serve as identifier, rank of producer according to index
               polymer_producer                             string, unique identifier, name of producer
               no_of_assets                                 numeric, metric, number of assets of the producer
               production_of_in_scope_polymers              numeric, metric in million metric tons, production of plolymers that are in-scope of preceding analysis
               flexible_format_contribution_to_sup_waste    numeric, metric in million metric tons, flexible form of contribution to sup waste
               rigid_format_contribution_to_sup_waste       numeric, metric in million metric tons, rigid form of contribution to sup waste
               total_contribution_to_sup_waste              numeric, metric in million metric tons, total contribution is the sum of flexible and rigid
  (i) missing values are in rows 346 and 347, we will just drop them, since there is no need for the dataset to be complete (actually it is incomplete by nature), and there are still       enough observations, also they seem to be missing at random 
  (i) no duplicated rows
  (i) no changes were made to data set
  (i) added a variable which gives the percentage of the sup_waste from the produced polymers
               
---
load packages
---
```{r load packages, setup}
library(tidyverse) # tidy data frame
library(lubridate) # functions to work with date-times and time-spans
library(scrubr) # like dplyr but specifically for occurrence data
```

---
import data
---
```{r}
df_raw <- read_csv(file = '../data/camera_dataset.csv')
df <- df_raw
```

---
first look at data
---
```{r}
head(df_raw, 25)
```
```{r}
tail(df_raw)
```
```{r}
summary(df_raw)
```

---
missing values
---
missing values are in rows 346 and 347, we will just drop them, since there is no need for the dataset to be complete (actually it is incomplete by nature), and there are still enough observations, also they seem to be missing at random 

```{r}
#create data frame with information on whether the value in each cell is missing
missing_by_column <- df_raw %>%
  is.na %>% # check is each cell is na
  as_tibble %>% # convert to data frame
  mutate(row_number = 1:nrow(.)) %>% # add a column with the row number
  gather(variable, is_missing, -row_number) # turn wide data into narrow data

# plot the missing values in our data frame
ggplot(missing_by_column, aes(x = variable, y = row_number, fill =  is_missing)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_grey(name = "",
                  labels = c("present", "missing")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=8)) +
  labs(x = "vairables in dataset",
       y = "rows / observations")
```
```{r}
# drop the rows with missing values 
df <- df[-c(346, 347), ]
```

---
duplicated rows
---
no duplicated rows

```{r}
# get row number of duplicated rows
duplicated_rows = tibble(duplicated = duplicated(df_raw), row = 1:nrow(df_raw)) %>%
  filter(duplicated == T)

# plot duplicated rows as black lines
ggplot(duplicated_rows, aes(xintercept = row)) +
  geom_vline(aes(xintercept = row)) + # plot a black line for each duplicated row
  ggtitle("indexes of duplicated rows") + # add title
  coord_flip() + # flip x and y axis
  scale_x_reverse() # reverse x axis
```

---
cleaning
---
nothing to clean

```{r}

```

---
additional variables
---
we can get the producer from the model name

```{r}

```

---
save processed data
---

no changes were made to data set





















