---
title: "describe univariate for camera data set"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/uni_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [X] describe & visualize single variables (univariate)
  - [X] gather interesting observations for further investigation
  - [X] gather possible new features for extraction
  
  
## insights 

  - brand column shows the count of cameras per brand, over 50% are produced by just 5 brands: Olympus, Sony, Canon, Kodak, Fujifilm
  - uni release_date: with increasing year the number of cameras in the data also increase, almost linear, this could be a sign of more cameras available on the market or just from the data acquisition 
  - comparing max_resolution and low_resolution shows that on median, the max resolution is higher, but both have a big spike around 2560 (resp 2592)
  - effective_pixels range between 0 and 13, mostly around 4 to 6, with two big spikes at 1 and 3
  - zoom_wide_W and zoom_tele_t are really far spread out but mostly centered around peaks, for zoom_wide_w its 28, 35, 38, for zoom_tele_t its 105a nd 114
  - normal_focus_range ranges mostly from 10 to 90 with only specific values and peaks at 30, 40, 50, 60, 90, but also rare values in between, macro_focus_range is centered around 7 with maximal 20 an modus 10, but quite some values in between compared to normal_focus_range 
  - uni storage_included: most data range from 1 to 48 mb with spikes at 8, 16, 32, but there are some big outliers, eg 450; not much information in this column except for technological progress and higher effective pixels require bigger storage, but these knowledge we already have from other variables, so lets drop this column in future
  - uni weight_inc_batteries: ranging from 100g to 600g with median 230g, and a lot of outliers up to 1800g, distribution is left skewed 
  - uni dimensions: ranging from 60 to 150 with median 100, also a big spike at 90, and some outliers up to 240 
  - uni price: range is 14 to 600, with median 200 and spikes at 120, 140, 220, and another isle at 400, outliers in the 1000 - 2000, but up to 8000, left skewed distribution
  
  - columns which can be dropped for future investigation: low_resolution, storage_included


## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(ggthemes) # for extra plot themes
library(plotly) # make ggplots interactive
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/camera_dataset_processed.csv') %>% column_to_rownames(var = 'model')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## univariate categorical brand

- brand column shows the count of cameras per brand, over 50% are produced by just 5 brands: Olympus, Sony, Canon, Kodak, Fujifilm

```{r}
# one variable, categorical x, show distribution
name = 'brand'
tmp_df <- df %>% rename(value = brand) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.85, 0.15), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, categorical x, 
# comparing the distribution of a variable against a hypothetical uniform distribution of that variable
name = c('camera count', 'brand share')
tmp_df <- df %>% rename(value = brand) %>% select(value)

library(gglorenz) #transformations for plotting Lorenz curve, https://github.com/jjchern/gglorenz

p1 <- tmp_df %>% 
  ggplot(aes(as.numeric(fct_infreq(value)))) +
    stat_lorenz(desc = FALSE) +
    coord_fixed() +
    geom_abline(linetype = 'dashed') +
    theme_minimal() +
    hrbrthemes::scale_x_percent() +
    hrbrthemes::scale_y_percent() +
    annotate_ineq(as.numeric(fct_infreq(tmp_df$value))) +
    ggtitle(paste("Lorenz curve for", name[1], sep=" ")) 
fig <- ggplotly(p1) %>% 
  layout(yaxis = list(title = paste("cumulative percentage of", name[1], sep=" ")), 
         xaxis = list(title = paste("cumulative percentage of", name[2], sep=" "))) 

fig
```


## univariate numeric release_date

- uni release_date: with increasing year the number of cameras in the data also increase, almost linear, this could be a sign of more cameras available on the market or just from the data acquisition  

```{r}
# one variable, continuous x, show distribution
name = 'release_date'
tmp_df <- df %>% rename(value = release_date) %>% select(value)

# https://ggplot2.tidyverse.org/reference/geom_dotplot.html
p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_density() +
    geom_histogram(binwidth = 1) +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, continuous x, show distribution
name = 'release_date'
tmp_df <- df %>% rename(value = release_date) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## compare univariate max_resolution and low_resolution

- comparing max_resolution and low_resolution shows that on median, the max resolution is higher, but both have a big spike around 2600

```{r}
# two variables, both continuous x, compare distributions
name = c('max_resolution', 'low_resolution')
tmp_df <- df %>% rename(max_res = max_resolution, low_res = low_resolution) %>%
  select(max_res, low_res) %>% pivot_longer(cols = c(max_res, low_res))

p1 <- tmp_df %>%
  ggplot(aes(x = value, fill = name)) +
    # geom_density(aes(colour = name), alpha = 0.5) +
    geom_histogram(binwidth = 100, alpha = 0.5, position = "identity") +
    theme_minimal() 
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = name, y = value, colour = name)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("compare ", name[1], "and", name[2], sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>%
  layout(xaxis = list(title = paste(name[1], "<br>", name[2], sep="")))

fig
```
```{r}
# one variable, continuous x, show distribution
name = c('max_resolution', 'low_resolution')
tmp_df <- df %>% rename(max_res = max_resolution, low_res = low_resolution) %>%
  select(max_res, low_res) %>% pivot_longer(cols = c(max_res, low_res)) %>% 
  add_count(value, name)

p1 <- tmp_df %>%
  ggplot(aes(x = value, fill = name, color = name)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    # stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity', trim = TRUE) +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = name, y = value, color = name)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("compare ", name[1], "and", name[2], sep=" "))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

p3 <- tmp_df %>%
  ggplot(aes(x = value, color = name)) +
    stat_density(geom="line", position = 'identity', trim = TRUE) + 
    theme_minimal()
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# https://plotly.com/r/subplots/
fig <- subplot(p3, p1, p2, nrows = 3, margin = 0, heights = c(0.1, 0.7, 0.2), shareX = TRUE) %>%
  layout(xaxis = list(title =name[1]))

fig
```



## univariate numeric effective_pixels

- effective_pixels range between 0 and 13, mostly around 4 to 6, with two big spikes at 1 and 3

```{r}
# one variable, continuous x, show distribution
name = 'effective_pixels'
tmp_df <- df %>% rename(value = effective_pixels) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_density() +
    geom_histogram(binwidth = 1) +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, continuous x, show distribution
name = 'effective_pixels'
tmp_df <- df %>% rename(value = effective_pixels) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## compare univariate zoom_wide_w and zoom_tele_t

- zoom_wide_W and zoom_tele_t are really far spread out but mostly centered around peaks, for zoom_wide_w its 28, 35, 38, for zoom_tele_t its 105a nd 114

```{r}
# two variables, both continuous x, compare distributions
name = c('zoom_wide_w', 'zoom_tele_t')
tmp_df <- df %>% rename(wide = zoom_wide_w, tele = zoom_tele_t) %>%
  select(wide, tele) %>% pivot_longer(cols = c(wide, tele))

p1 <- tmp_df %>%
  ggplot(aes(x = value, fill = name)) +
    # geom_density(aes(colour = name), alpha = 0.5) +
    geom_histogram(binwidth = 1, alpha = 0.5, position = "identity") +
    theme_minimal() 
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = name, y = value, colour = name)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("compare ", name[1], "and", name[2], sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>%
  layout(xaxis = list(title = paste(name[1], "<br>", name[2], sep="")))

fig
```
```{r}
# one variable, continuous x, show distribution
name = c('zoom_wide_w', 'zoom_tele_t')
tmp_df <- df %>% rename(wide = zoom_wide_w, tele = zoom_tele_t) %>%
  select(wide, tele) %>% pivot_longer(cols = c(wide, tele)) %>% 
  add_count(name, value)

p1 <- tmp_df %>%
  ggplot(aes(x = value, fill = name, color = name)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    # stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity', trim = TRUE) +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = name, y = value, color = name)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("compare ", name[1], "and", name[2], sep=" "))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

p3 <- tmp_df %>%
  ggplot(aes(x = value, color = name)) +
    stat_density(geom="line", position = 'identity', trim = TRUE) + 
    theme_minimal()
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# https://plotly.com/r/subplots/
fig <- subplot(p3, p1, p2, nrows = 3, margin = 0, heights = c(0.1, 0.7, 0.2), shareX = TRUE) %>%
  layout(xaxis = list(title =name[1]))

fig
```


## compare univariate normal_focus_range and macro_focus_range

- normal_focus_range ranges mostly from 10 to 90 with only specific values and peaks at 30, 40, 50, 60, 90, but also rare values in between, macro_focus_range is centered around 7 with maximal 20 an modus 10, but quite some values in between compared to normal_focus_range 

```{r}
# two variables, both continuous x, compare distributions
name = c('normal_focus_range', 'macro_focus_range')
tmp_df <- df %>% rename(normal = normal_focus_range, macro = macro_focus_range) %>%
  select(normal, macro) %>% pivot_longer(cols = c(normal, macro))

p1 <- tmp_df %>%
  ggplot(aes(x = value, fill = name)) +
    # geom_density(aes(colour = name), alpha = 0.5) +
    geom_histogram(binwidth = 1, alpha = 0.5, position = "identity") +
    theme_minimal() 
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = name, y = value, colour = name)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("compare ", name[1], "and", name[2], sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>%
  layout(xaxis = list(title = paste(name[1], "<br>", name[2], sep="")))

fig
```
```{r}
# one variable, continuous x, show distribution
name = c('normal_focus_range', 'macro_focus_range')
tmp_df <- df %>% rename(normal = normal_focus_range, macro = macro_focus_range) %>%
  select(normal, macro) %>% pivot_longer(cols = c(normal, macro)) %>% 
  add_count(name, value)

p1 <- tmp_df %>%
  ggplot(aes(x = value, fill = name, color = name)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    # stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity', trim = TRUE) +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = name, y = value, color = name)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("compare ", name[1], "and", name[2], sep=" "))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

p3 <- tmp_df %>%
  ggplot(aes(x = value, color = name)) +
    stat_density(geom="line", position = 'identity', trim = TRUE) + 
    theme_minimal()
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# https://plotly.com/r/subplots/
fig <- subplot(p3, p1, p2, nrows = 3, margin = 0, heights = c(0.1, 0.7, 0.2), shareX = TRUE) %>%
  layout(xaxis = list(title =name[1]))

fig
```


## univariate numeric storage_included

- uni storage_included: most data range from 1 to 48 mb with spikes at 8, 16, 32, but there are some big outliers, eg 450; not much information in this column except for technological progress and higher effective pixels require bigger storage, but these knowledge we already have from other variables, so lets drop this column in future

```{r}
# one variable, continuous x, show distribution
name = 'storage_included'
tmp_df <- df %>% rename(value = storage_included) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_density() +
    geom_histogram(binwidth = 1) +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, continuous x, show distribution
name = 'storage_included'
tmp_df <- df %>% rename(value = storage_included) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric weight_inc_batteries

- uni weight_inc_batteries: ranging from 100g to 600g with median 230g, and a lot of outliers up to 1800g, distribution is left skewed 

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_density() +
    geom_histogram(binwidth = 10) +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric dimensions

- uni dimensions: ranging from 60 to 150 with median 100, also a big spike at 90, and some outliers up to 240 

```{r}
# one variable, continuous x, show distribution
name = 'dimensions'
tmp_df <- df %>% rename(value = dimensions) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_density() +
    geom_histogram(binwidth = 10) +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, continuous x, show distribution
name = 'dimensions'
tmp_df <- df %>% rename(value = dimensions) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric price

- uni price: range is 14 to 600, with median 200 and spikes at 120, 140, 220, and another isle at 400, outliers in the 1000 - 2000, but up to 8000, left skewed distribution

```{r}
# one variable, continuous x, show distribution
name = 'price'
tmp_df <- df %>% rename(value = price) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_density() +
    geom_histogram(binwidth = 20) +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```
```{r}
# one variable, continuous x, show distribution
name = 'price'
tmp_df <- df %>% rename(value = price) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```




## experimental plot area

- https://cran.r-project.org/web/packages/ggbeeswarm/vignettes/usageExamples.pdf
- https://stackoverflow.com/questions/32533356/is-there-a-way-to-make-the-density-function-in-r-use-counts-vs-probability

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value)

tmp_sum_n <- tmp_df %>% distinct(value, .keep_all = TRUE) %>% select(n)
p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_histogram(aes(y = ..density..), binwidth = 10) +
    # geom_point(aes(y = n), alpha = 0.2) +
    geom_spoke(aes(y = 0, angle = pi/2, radius = n), alpha = 0.5, lwd = 1, stat = "unique") +
    stat_density(aes(y = ..density..* nrow(tmp_df %>% distinct) * density(tmp_df$value, na.rm = TRUE)$bw), geom = 'line') +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value)

p3 <- tmp_df %>%
  distinct(value) %>%
  ggplot(aes(x = value)) +
    stat_density(geom = 'line') + 
    theme_minimal()
p3 <- ggplotly(p3) %>% layout()

p3
```

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value)

library(ggbeeswarm)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_quasirandom(aes(y = 0), groupOnX=FALSE, width = max(tmp_df$n), alpha = 0.25, bandwidth = 1/length(unique(tmp_df$value))) +
    stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line') +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
  # ..density.. * sum(tmp_df$n) or ..scaled.. * max(tmp_df$n) or ..density.. * nrow(tmp_df) * density(tmp_df$value)$bw or ..count..
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value) %>% na.omit()

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
(max(tmp_df$value)-min(tmp_df$value))
```

```{r}
# one variable, continuous x, show distribution
name = 'weight_inc_batteries'
tmp_df <- df %>% rename(value = weight_inc_batteries) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_point(aes(y = n), alpha = 0.5, stat = 'unique') +
    stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


```{r}
# dat <- c(rnorm(100), rnorm(100,5))
# dat <- c(rbinom(500, 100, 0.4), rpois(500, 25))
# dat <- c(rbinom(500, 100, 0.4), rnorm(1000,5))
dat <- c(rbinom(100, 10, 0.3), rnorm(100,5), rep(10, 16))


name = 'weight_inc_batteries'
tmp_df <- as.tibble(dat) %>% rename(value = value) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    # geom_histogram(aes(y = ..density..), binwidth = 10) +
    # geom_point(aes(y = n), alpha = 0.2) +
    geom_spoke(aes(y = -n, angle = pi/2, radius = 2*n), 
               alpha = 0.3, lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y = ..scaled.. * max(tmp_df$n)), geom = 'line') + # or ..scaled.. * max(tmp_df$n)
    theme_minimal()  
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


