---
title: "cleaning for camera data set"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/clean_",
  warning = FALSE,
  message=FALSE
)
```



## purpose of notebook

  - [x] initial look at data to get a basic understanding and gather todos -> df_raw
  - [x] tidy up and cleaning of data set -> df
  - [x] all processing of the data set happens here and is saved to pre-/processed csv-> df
 

## observations

  - column information:  
  
  |  # | attribute | description & properties | missing values | dependent? |                                                                                              
  |:--:|:---------:|:-----------------------------------------------------------:|:---:|:---:|
  | 1 | Model | brand and model of the camera, chr, unique, can serve as identifier | non | no |
  | 2 | Release date | release date of camera, time series and count, ordered, numerical, discrete | non | no |
  | 3 | Max resolution | numerical, metric in number of pixel columns (width), maximal image resolution | almost non | no |
  | 4 | Low resolution | numerical, metric in number of pixel columns (width), standard image resolution | some, not MAR, there is only max option | yes, capped by max resolution |
  | 5 | Effective pixels | numerical, metric in megapixels, arbitrarily rounded, count of image pixel contributing to the final image | no | yes, directly by max resolution |
  | 6 | Zoom wide (W) | numerical, metric in mm, focal lengths that create a wide field of view beyond 35mm | some, not MAR, professional cameras don't include lens | yes, by lens |
  | 7 | Zoom tele (T) | numerical, metric in mm, focal lengths that create a narrow field of view beyond 80mm | some, not MAR, professional cameras don't include lens | yes, by lens |
  | 8 | Normal focus range | numerical, metric in mm, minimum distance at which focus can be obtained, feature of lens | some, some MAR, professional cameras don't include lens | yes, by lens |
  | 9 | Macro focus range | numerical, metric in mm, reduced focus range (or working range) to focus on close objects | some, some MAR, professional cameras don't include lens | yes, by lens |
  | 10 | Storage included | numerical, metric in megabyte, storage of included SD-card | a few, not MAR, SD-card just not included | no |  
  | 11 | Weight (inc. batteries) | numerical, metric in g, weight of the camera including batteries | a few, MAR | no |
  | 12 | Dimensions | numerical, metric in mm, width of the camera dimensions | a few, MAR | no |
  | 13 | Price | numerical, metric in unknown currency, price when released | non | no |
  
  - missing values with NA are in rows 346 and 347, but there are a lot more 0 which are mostly not missing at random (MAR), but some times they do, later we will replace the 0 with        NA, but for investigation not drop or impute them
  - no duplicated rows
  - cleaning: 
    + replace 0 with NA, since these entries would disturb the picture of the data set otherwise, except for Effective pixels column
    + get clean column names with janitor package, without () and white space
    + column Model serves as unique identifier, and thus can be set as row name (number)
  - added variables: 
    + we can get the brand from the model name as a new variable (categorical, no order)
               

## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(lubridate) # functions to work with date-times and time-spans
library(scrubr) # like dplyr but specifically for occurrence data
library(janitor) # expedite the initial data exploration and cleaning that comes with any new data set
```


## import data

```{r}
df_raw <- read_csv(file = '../data/camera_dataset.csv')
df <- df_raw
```


## first look at data

```{r}
head(df_raw, 25)
```
```{r}
tail(df_raw)
```
```{r}
summary(df_raw)
```


## missing values

- missing values with NA are in rows 346 and 347, but there are a lot more 0 which are mostly not missing at random (MAR), but some times they do, later we will replace the 0 with NA, but for investigation not drop or impute them

```{r fig.height=20, fig.width=10}
#create data frame with information on whether the value in each cell is zero
tmp_df <- df_raw == 0
missing_by_column <- tmp_df %>%
  as_tibble %>% # convert to data frame
  mutate(row_number = 1:nrow(.)) %>% # add a column with the row number
  gather(variable, is_missing, -row_number) # turn wide data into narrow data

# plot the missing values in our data frame
ggplot(missing_by_column, aes(x = variable, y = row_number, fill =  is_missing)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_grey(name = "",
                  labels = c("present", "is zero", "is NA")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10)) +
  labs(x = "vairables in dataset",
       y = "rows / observations")
```


## duplicated rows

- no duplicated rows

```{r}
# get row number of duplicated rows
duplicated_rows <- tibble(duplicated = duplicated(df_raw), row = 1:nrow(df_raw)) %>%
  filter(duplicated == T)

# plot duplicated rows as black lines
ggplot(duplicated_rows, aes(xintercept = row)) +
  geom_vline(aes(xintercept = row)) + # plot a black line for each duplicated row
  ggtitle("indexes of duplicated rows") + # add title
  coord_flip() + # flip x and y axis
  scale_x_reverse() # reverse x axis
```


## cleaning

- replace 0 with NA, since these entries would disturb the picture of the data set otherwise, except for Effective pixels column  
- get clean column names with janitor package, without () and white space  
- column Model serves as unique identifier, and thus can be set as row name (number)  

```{r}
df[df == 0] <- NA # replace all 0 with NA

df <- df %>%
  mutate(`Effective pixels` = df_raw$`Effective pixels`) %>% # bring back the zeros for Effective pixels column
  clean_names() %>% # get clean column names with janitor package
  column_to_rownames(var = 'model') # set model columns as the row name
```


## additional variables

- we can get the brand from the model name as a new variable (categorical, no order)

```{r}
df <- df %>%
  mutate(brand = str_extract(df_raw$Model,"(\\w+)")) %>% # extract the first word of the model column as brand name
  relocate(brand) # relocate the brand column at the first position in the df
```

```{r}
# check results from cleaning and added variables
head(df)
```


## save processed data

```{r}
df %>% rownames_to_column(var = 'model') %>% write_csv(file = '../data/camera_dataset_processed.csv')
```

