---
title: "describe bivariate for camera data set"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/multi_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] describe & visualize relationship between variables (multivariate)
  - [ ] gather interesting observations for further investigation
  - [ ] gather possible new features for extraction
  
  
## insights 

  - 


## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(ggthemes) # for extra plot themes
library(plotly) # make ggplots interactive
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/camera_dataset_processed.csv') %>% column_to_rownames(var = 'model')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## bivariate numeric max_resolution over categorical brand 

- 

```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('max_resolution', 'brand')
tmp_df <- df %>% rename(x = brand, y = max_resolution) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('max_resolution', 'brand')
tmp_df <- df %>% rename(x = brand, y = max_resolution) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric effective_pixels over categorical brand 

- 

```{r fig.height=12, fig.width=10}
# two variables, continuous x, categorical y, show trend and distribution
name = c('effective_pixels', 'brand')
tmp_df <- df %>% rename(x = brand, y = effective_pixels) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('effective_pixels', 'brand')
tmp_df <- df %>% rename(x = brand, y = effective_pixels) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric weight_inc_batteries over categorical brand

-

```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('weight_inc_batteries', 'brand')
tmp_df <- df %>% rename(x = brand, y = weight_inc_batteries) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('weight_inc_batteries', 'brand')
tmp_df <- df %>% rename(x = brand, y = weight_inc_batteries) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric price over categorical brand

-

```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('price', 'brand')
tmp_df <- df %>% rename(x = brand, y = price) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('price', 'brand')
tmp_df <- df %>% rename(x = brand, y = price) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric max_resolution over numeric weight_inc_batteries

-

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('weight_inc_batteries', 'max_resolution')
tmp_df <- df %>% rename(x = weight_inc_batteries, y = max_resolution) %>% select(x, y)

# https://ggplot2.tidyverse.org/reference/geom_smooth.html
point_plot <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(alpha = 0.5, size = 1) +
    geom_rug(alpha = 0.5) + # two 1d marginal distributions, display individual cases so are best used with smaller datasets
    geom_density_2d(alpha = 0.2, bins = 4) +# 2D kernel density estimation using MASS::kde2d() and display the results with contours
    geom_smooth(fill = "grey90") + # aids the eye in seeing patterns in the presence of overplotting
    geom_point(alpha = 0.75) + # point geom is used to create scatterplots
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
point_plot <- ggplotly(point_plot) %>% layout(xaxis = list(showticklabels = FALSE))

x_density_plot <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + # draws kernel density estimate, which is a smoothed version of the histogram
    # geom_histogram(binwidth = 1) +
    theme_minimal()
x_density_plot <- ggplotly(x_density_plot) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE), 
                                                      xaxis = list(showticklabels = FALSE, showgrid = FALSE))

y_density_plot <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + # draws kernel density estimate, which is a smoothed version of the histogram
    # geom_histogram(binwidth = 1) +
    coord_flip() +
    theme_minimal()
y_density_plot <- ggplotly(y_density_plot) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE), 
                                                      xaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://ggplot2.tidyverse.org/reference/geom_quantile.html
qualtile_plot <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_quantile(alpha = 0.8) + # fits a quantile regression to the data and draws the fitted quantiles with lines
    theme_minimal()
qualtile_plot <- ggplotly(qualtile_plot) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
sub1 <- subplot(x_density_plot, plotly_empty(), point_plot, y_density_plot, nrows = 2, margin = 0, 
                heights = c(0.15, 0.85), widths = c(0.9, 0.1), shareX = TRUE, shareY = TRUE, titleX = FALSE, titleY = FALSE) %>% 
  layout()
sub2 <- subplot(qualtile_plot, plotly_empty(), margin = 0, widths = c(0.9, 0.1), titleX = FALSE, titleY = FALSE) %>% layout()
fig <- subplot(sub1, sub2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('weight_inc_batteries', 'max_resolution')
tmp_df <- df %>% rename(x = weight_inc_batteries, y = max_resolution) %>% select(x, y) %>% add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    geom_density2d(alpha = 0.5, bins = 4) +
    # geom_quantile(alpha = 0.4) +
    # geom_smooth(fill = "grey90") +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```


## bivariate numeric max_resolution over numeric weight_inc_batteries

-

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('low_resolution', 'max_resolution')
tmp_df <- df %>% rename(x = low_resolution, y = max_resolution) %>% select(x, y) %>% add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    geom_density2d(alpha = 0.5, bins = 4) +
    # geom_quantile(alpha = 0.4) +
    # geom_smooth(fill = "grey90") +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```


## bivariate numeric max_resolution over numeric price

-

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('price', 'max_resolution')
tmp_df <- df %>% rename(x = price, y = max_resolution) %>% select(x, y) %>% add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    geom_density2d(alpha = 0.5, bins = 4) +
    # geom_quantile(alpha = 0.4) +
    # geom_smooth(fill = "grey90") +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```


## bivariate numeric normal_focus_range over numeric macro_focus_range

-

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('normal_focus_range', 'macro_focus_range')
tmp_df <- df %>% rename(x = normal_focus_range, y = macro_focus_range) %>% select(x, y) %>% add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    geom_density2d(alpha = 0.5, bins = 4) +
    # geom_quantile(alpha = 0.4) +
    # geom_smooth(fill = "grey90") +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```





