---
title: "summary for camera data set"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/summary_",
  warning = FALSE,
  message=FALSE
)
```



## purpose of notebook

  - [ ] summarize all insights and ideas from the other notebooks, as well as good exploratory plots
   
   
## information

found on: https://www.kaggle.com/crawford/1000-cameras-dataset/
  
with acknowledgements to:  
    These datasets have been gathered and cleaned up by Petra Isenberg, Pierre Dragicevic and Yvonne Jansen. The original source can be found here
    https://perso.telecom-paristech.fr/eagan/class/igr204/datasets
  
licensed under: Creative Commons Attribution-ShareAlike 3.0 License https://creativecommons.org/licenses/by-sa/3.0/
  
Context: Some camera enthusiast went and described 1,000 cameras based on 13 properties!
   
    
## domain information 

Content: The 13 properties of each camera  
   1. Model  
   2. Release date  
   3. Max resolution  
   4. Low resolution  
   5. Effective pixels  
   6. Zoom wide (W)  
   7. Zoom tele (T)  
   8. Normal focus range  
   9. Macro focus range  
   10. Storage included  
   11. Weight (inc. batteries)  
   12. Dimensions  
   13. Price  

- https://en.wikipedia.org/wiki/Image_resolution
- https://www.adorama.com/alc/wide-angle-vs-telephoto-which-lens-should-you-choose/
- https://photo.stackexchange.com/questions/86323/what-does-it-mean-when-macro-focus-range-and-normal-focus-range-are-equal

  
## summary highlights
  



## stories


 
 
## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(plotly) # make ggplots interactive
```


## import data

```{r}
df <- read_csv(file = '../data/camera_dataset_processed.csv') %>% column_to_rownames(var = 'model')
```

## overview

```{r}
head(df)
```

```{r}
summary(df)
```


## observations from clean nb




## insights from describe uni

  


## insights from describe multi



