---
title: "describe univariate for camera data set"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/multi_",
  warning = FALSE
)
```



## purpose of notebook

  - [ ] describe & visualize relationship between variables (multivariate)
  - [ ] gather interesting observations for further investigation
  - [ ] gather possible new features for extraction
  
  
## insights 

  - 


## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(ggthemes) # for extra plot themes
library(plotly) # make ggplots interactive
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/camera_dataset_processed.csv') %>% column_to_rownames(var = 'model')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## bivariate numeric max_resolution over categorical brand 

- 

```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('max_resolution', 'brand')
tmp_df <- df %>% rename(x = brand, y = max_resolution) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('max_resolution', 'brand')
tmp_df <- df %>% rename(x = brand, y = max_resolution) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric effective_pixels over categorical brand 

- 

```{r fig.height=12, fig.width=10}
# two variables, continuous x, categorical y, show trend and distribution
name = c('effective_pixels', 'brand')
tmp_df <- df %>% rename(x = brand, y = effective_pixels) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('effective_pixels', 'brand')
tmp_df <- df %>% rename(x = brand, y = effective_pixels) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric weight_inc_batteries over categorical brand

-

```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('weight_inc_batteries', 'brand')
tmp_df <- df %>% rename(x = brand, y = weight_inc_batteries) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('weight_inc_batteries', 'brand')
tmp_df <- df %>% rename(x = brand, y = weight_inc_batteries) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric price over categorical brand

-

```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('price', 'brand')
tmp_df <- df %>% rename(x = brand, y = price) %>% select(x, y) %>% mutate(x = fct_infreq(x))

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    # geom_jitter(aes(color = x), alpha = I(0.5), width = 0.2, height = 0.1) + # for continuous x
    geom_count(aes(color = x), alpha = I(0.5)) +  # for discrete x
    geom_boxplot(fill=NA, lwd = 0.25) +
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```
```{r fig.height=12, fig.width=10}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('price', 'brand')
tmp_df <- df %>% rename(x = brand, y = price) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), lwd = 1, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric ... over numeric ...

-

```{r}
# # two variables, continuous x, continuous y, show trend and distribution
# name = c('rank', 'no_of_assets')
# df <- plastic %>% rename(x = rank, y = no_of_assets) %>% select(x, y) %>% mutate(x = -x) # x = -x to show rank as increasing from low to high 
# 
# # https://ggplot2.tidyverse.org/reference/geom_smooth.html
# point_plot <- df %>%
#   ggplot(aes(x = x, y = y)) +
#     # geom_jitter(alpha = 0.5, size = 1) +
#     geom_rug(alpha = 0.5) + # two 1d marginal distributions, display individual cases so are best used with smaller datasets
#     geom_density_2d(alpha = 0.2, bins = 4) +# 2D kernel density estimation using MASS::kde2d() and display the results with contours
#     geom_smooth(fill = "grey90") + # aids the eye in seeing patterns in the presence of overplotting
#     geom_point(alpha = 0.75) + # point geom is used to create scatterplots
#     theme_minimal() +
#     ggtitle(paste("trend of", name[2], "over", name[1], sep=" ")) 
# point_plot <- ggplotly(point_plot) %>% layout(xaxis = list(showticklabels = FALSE))
# 
# x_density_plot <- df %>%
#   ggplot(aes(x = x)) +
#     stat_density(geom="line") + # draws kernel density estimate, which is a smoothed version of the histogram
#     # geom_histogram(binwidth = 1) +
#     theme_minimal() 
# x_density_plot <- ggplotly(x_density_plot) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE), xaxis = list(showticklabels = FALSE, showgrid = FALSE))
# 
# y_density_plot <- df %>%
#   ggplot(aes(x = y)) +
#     stat_density(geom="line") + # draws kernel density estimate, which is a smoothed version of the histogram
#     # geom_histogram(binwidth = 1) +
#     coord_flip() +
#     theme_minimal() 
# y_density_plot <- ggplotly(y_density_plot) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE), xaxis = list(showticklabels = FALSE, showgrid = FALSE))
# 
# # https://ggplot2.tidyverse.org/reference/geom_quantile.html
# qualtile_plot <- df %>%
#   ggplot(aes(x = x, y = y)) +
#     geom_quantile(alpha = 0.8) + # fits a quantile regression to the data and draws the fitted quantiles with lines
#     theme_minimal() 
# qualtile_plot <- ggplotly(qualtile_plot) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))
# 
# # merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
# sub1 <- subplot(x_density_plot, plotly_empty(), point_plot, y_density_plot, nrows = 2, margin = 0, heights = c(0.15, 0.85), widths = c(0.9, 0.1), shareX = TRUE, shareY = TRUE, titleX = FALSE, titleY = FALSE) %>% layout()
# sub2 <- subplot(qualtile_plot, plotly_empty(), margin = 0, widths = c(0.9, 0.1), titleX = FALSE, titleY = FALSE) %>% layout()
# fig <- subplot(sub1, sub2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))
#   
# fig
```







