---
title: "describe univariate for Estimation of global male circumcision"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/uni_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] describe & visualize single variables (univariate)
  - [ ] gather interesting observations for further investigation
  - [ ] gather possible new features for extraction
  
  
## insights 

  - male_circumcision_percent: it seems that there are two populations of countries, one of mostly circumized males, the other of mostly not-circumzied males; but also the few countries in the middle around 50% are interesting 
  

## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(plotly) # make ggplots interactive
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/male_circumcision.csv')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## univariate categorical region

- region: 

```{r}
# one variable, categorical x, show distribution
name = 'region'
tmp_df <- df %>% rename(value = region) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical sub_region

- sub_region: 

```{r}
# one variable, categorical x, show distribution
name = 'sub_region'
tmp_df <- df %>% rename(value = sub_region) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical intermediate_region

- intermediate_region: 

```{r}
# one variable, categorical x, show distribution
name = 'intermediate_region'
tmp_df <- df %>% rename(value = intermediate_region) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric male_circumcision_percent

- male_circumcision_percent: it seems that there are two populations of countries, one of mostly circumized males, the other of mostly not-circumzied males; but also the few countries in the middle around 50% are interesting  

```{r}
# one variable, continuous x, show distribution
name = 'male_circumcision_percent'
tmp_df <- df %>% rename(value = male_circumcision_percent) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric cia_total_males

- cia_total_males:

```{r}
# one variable, continuous x, show distribution
name = 'cia_total_males'
tmp_df <- df %>% rename(value = cia_total_males) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate numeric total_males_circumcized

- total_males_circumcized:

```{r}
# one variable, continuous x, show distribution
name = 'total_males_circumcized'
tmp_df <- df %>% rename(value = total_males_circumcized) %>% select(value) %>% add_count(value)

p1 <- tmp_df %>%
  ggplot(aes(x = value)) +
    geom_spoke(aes(y = -n, radius = 2*n, angle = pi/2, text = paste0("value: ", value, "\ncount: ", n)),
               alpha = 0.5, stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    stat_density(aes(y =..scaled.. * max(tmp_df$n) * 1.5), geom = 'line', position = 'identity') +
    theme_minimal()  
p1 <- ggplotly(p1, tooltip = 'text') %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = value)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) 
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.8, 0.2), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```


## univariate categorical basis

- basis: 

```{r}
# one variable, categorical x, show distribution
name = 'basis'
tmp_df <- df %>% rename(value = basis) %>% select(value)

p1 <- tmp_df %>%
  ggplot(aes(x = fct_infreq(value))) +
    geom_bar() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p1 <- ggplotly(p1) %>% layout()

p2 <- tmp_df %>%
  ggplot(aes(x = 1, y = fct_infreq(value))) +
    geom_boxplot(fill = 'lightgrey') +
    theme_minimal() +
    coord_flip() +
    ggtitle(paste("distribution of", name, sep=" ")) +
    theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10))
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE))

# https://plotly.com/r/subplots/
fig <- subplot(p1, p2, nrows = 2, margin = 0, heights = c(0.9, 0.1), shareX = TRUE) %>% 
  layout(xaxis = list(title = name))

fig
```




