---
title: "describe bivariate with time for New York City Restaurant Inspection Results"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/time_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] describe & visualize relationship between variables, one of them time-like
  - [ ] gather interesting observations for further investigation
  - [ ] gather possible new features for extraction
  
  
## insights 

  - 


## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame and so much more
library(plotly) # make ggplots interactive
library(lubridate) # functions to work with date-times and time-spans
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/restaurants_data_processed.csv')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## univariate time inspection_date over categorical boro

- 

```{r}
# https://r4ds.had.co.nz/dates-and-times.html
# Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.
name = c('boro', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = boro) %>% select(t, y) %>% 
  filter(t != '1900-01-01')

fig <- tmp_df %>%
  ggplot(aes(x = t, col = y)) +
    geom_freqpoly(bins = 30) + # binwidth = 86400 seconds = 1 day, 600 s = 10 minutes
    theme_minimal() +
    ggtitle("frequency plot for StartTime") +
    facet_grid(rows = vars(y))

ggplotly(fig)
```


## univariate time inspection_date over categorical boro - wday

- 

```{r}
# https://r4ds.had.co.nz/dates-and-times.html
# Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.
name = c('boro', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = boro) %>% select(t, y) %>% filter(t != '1900-01-01') %>%
  mutate(t = wday(t, label = TRUE))

fig <- tmp_df %>%
  ggplot(aes(x = t, fill = y)) +
    geom_bar() +
    facet_grid(rows = vars(y)) +
    theme_minimal() +
    ggtitle("distribution over week days")

ggplotly(fig)
```


## univariate time inspection_date over categorical boro - year

- 

```{r}
# https://r4ds.had.co.nz/dates-and-times.html
# Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.
name = c('boro', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = boro) %>% select(t, y) %>% filter(t != '1900-01-01') %>%
  mutate(t = year(t))

fig <- tmp_df %>%
  ggplot(aes(x = t, fill = y)) +
    geom_bar() +
    facet_grid(rows = vars(y)) +
    theme_minimal() +
    ggtitle("distribution over years")

ggplotly(fig)
```


## univariate time inspection_date over categorical boro - month

- 

```{r}
# https://r4ds.had.co.nz/dates-and-times.html
# Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.
name = c('boro', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = boro) %>% select(t, y) %>% filter(t != '1900-01-01') %>%
  mutate(t = month(t, label = TRUE))

fig <- tmp_df %>%
  ggplot(aes(x = t, fill = y)) +
    geom_bar() +
    facet_grid(rows = vars(y)) +
    theme_minimal() +
    ggtitle("distribution over months")

ggplotly(fig)
```


## univariate time inspection_date over categorical boro - mday

- 

```{r}
# https://r4ds.had.co.nz/dates-and-times.html
# Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.
name = c('boro', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = boro) %>% select(t, y) %>% filter(t != '1900-01-01') %>%
  mutate(t = mday(t))

fig <- tmp_df %>%
  ggplot(aes(x = t, fill = y)) +
    geom_bar() +
    facet_grid(rows = vars(y)) +
    theme_minimal() +
    ggtitle("distribution over month days")

ggplotly(fig)
```


## univariate time inspection_date over categorical boro - yday

- 

```{r}
# https://r4ds.had.co.nz/dates-and-times.html
# Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.
name = c('boro', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = boro) %>% select(t, y) %>% filter(t != '1900-01-01') %>%
  mutate(t = yday(t))

fig <- tmp_df %>%
  ggplot(aes(x = t, fill = y)) +
    geom_bar() +
    facet_grid(rows = vars(y)) +
    theme_minimal() +
    ggtitle("distribution over year days")

ggplotly(fig)
```


## univariate numeric score over time inspection_date

- 

```{r}
# two variables, continuous/discrete x, time t, show trend and distribution
name = c('score', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = score) %>% select(t, y) %>% 
  add_count(t, y)


so.q2 <- function(y){ quantile(y, na.rm = TRUE)[2] }
so.q4 <- function(y){ quantile(y, na.rm = TRUE)[4] }

p1 <- tmp_df %>%
  ggplot(aes(x = t, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    stat_summary(fun.y = so.q2, geom = 'line', size = 0.25) + 
    stat_summary(fun.y = so.q4, geom = 'line', size = 0.25) + 
    # stat_summary(fun.y = 'median', geom = 'line', size = 0.25) + 
    # geom_spoke(aes(x = t + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi),
               # alpha = I(0.5), stat = 'unique', color = 'darkgreen') + 
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "over time", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[2]), yaxis = list(title = name[1]))

fig
```


## univariate numeric score over time inspection_date -  month

- 

```{r}
# two variables, continuous/discrete x, time t, show trend and distribution
name = c('score', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = score) %>% select(t, y) %>% 
  mutate(t = month(t)) %>% add_count(t, y)


so.q2 <- function(y){ quantile(y, na.rm = TRUE)[2] }
so.q4 <- function(y){ quantile(y, na.rm = TRUE)[4] }

p1 <- tmp_df %>%
  ggplot(aes(x = t, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    stat_summary(fun.y = so.q2, geom = 'line', size = 0.25) + 
    stat_summary(fun.y = so.q4, geom = 'line', size = 0.25) + 
    # stat_summary(fun.y = 'median', geom = 'line', size = 0.25) + 
    geom_spoke(aes(x = t + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi),
    alpha = I(0.5), stat = 'unique', color = 'darkgreen') +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "over time", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[2]), yaxis = list(title = name[1]))

fig
```


## univariate numeric score over time inspection_date -  wday

- 

```{r}
# two variables, continuous/discrete x, time t, show trend and distribution
name = c('score', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, y = score) %>% select(t, y) %>% 
  mutate(t = wday(t)) %>% add_count(t, y)


so.q2 <- function(y){ quantile(y, na.rm = TRUE)[2] }
so.q4 <- function(y){ quantile(y, na.rm = TRUE)[4] }

p1 <- tmp_df %>%
  ggplot(aes(x = t, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    stat_summary(fun.y = so.q2, geom = 'line', size = 0.25) + 
    stat_summary(fun.y = so.q4, geom = 'line', size = 0.25) + 
    # stat_summary(fun.y = 'median', geom = 'line', size = 0.25) + 
    geom_spoke(aes(x = t + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi),
    alpha = I(0.5), stat = 'unique', color = 'darkgreen') +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "over time", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[2]), yaxis = list(title = name[1]))

fig
```


## bivariate numeric score over numeric camis over time result

- nothing to see here. move along

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('score', 'camis', 'inspection_date')
tmp_df <- df %>% rename(t = inspection_date, x = score, y = camis) %>% select(t, x, y) %>%  
  add_count(x, y) %>% arrange(t)


p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_segment(aes(xend = c(tail(x, n=-1), NA), yend = c(tail(y, n=-1), NA), color = t), lwd=0.4, alpha=0.7) +
    viridis::scale_color_viridis(name = name[3], option = 'H') +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```


