---
title: "describe bivariate for New York City Restaurant Inspection Results"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/multi_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] describe & visualize relationship between variables (multivariate)
  - [ ] gather interesting observations for further investigation
  - [ ] gather possible new features for extraction
  
  
## insights 

  - 


## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(plotly) # make ggplots interactive
```


## import data

```{r, message=FALSE}
df <- read_csv(file = '../data/restaurants_data_processed.csv')
```


## overview

```{r}
head(df)
```
```{r}
summary(df)
```


## scagnostics

- 

```{r}
name = c('')
tmp_df <- df %>% select(-camis, -dba, -violation_description, -latitude, -longitude, -street, -grade, -grade_date) %>% na.omit()

library(scagnostics) # Scagnostics (scatterplot diagnostics) summarize potentially interesting patterns in 2d scatterplots.

scagnostics_results <- scagnostics(tmp_df)
head(scagnostics_results)

e <- scagnosticsExemplars(scagnostics_results)
o <- scagnosticsOutliers(scagnostics_results)
g <- scagnosticsGrid(scagnostics_results)

# look at outliers
o[o]
go <- g[o,]
par(mfrow = c(2,2))
plot(tmp_df[[go$x]], tmp_df[[go$y]], pch=19, 
     xlab=names(tmp_df)[go$x], ylab=names(tmp_df)[go$y])

# find and plot exemplars
e[e]
ge <- g[e,]
par(mfrow = c(2,2))
for(i in 1:dim(ge)[1])
  plot(tmp_df[[ge$x[i]]], tmp_df[[ge$y[i]]], pch=19,
       xlab=names(tmp_df)[ge$x[i]], ylab=names(tmp_df)[ge$y[i]])

par(mfrow = c(2,2))
for(i in 1:dim(g)[1])
  plot(tmp_df[[g$x[i]]], tmp_df[[g$y[i]]], pch=19,
       xlab=names(tmp_df)[g$x[i]], ylab=names(tmp_df)[g$y[i]])
```


## bivariate numeric latitude over numeric longitude

- lat over long: actually this should be a geospatial plot

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('latitude', 'longitude')
tmp_df <- df %>% rename(x = latitude, y = longitude) %>% select(x, y) %>% add_count(x, y) %>% filter(x != 0)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    geom_density2d(alpha = 0.5, bins = 4) +
    # geom_quantile(alpha = 0.4) +
    # geom_smooth(fill = "grey90") +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```


## bivariate numeric score over categorical boro 

- 

```{r}
# two variables, continuous/discrete x, categorical y, show trend and distribution
name = c('score', 'boro')
tmp_df <- df %>% rename(x = boro, y = score) %>% select(x, y) %>% mutate(x = fct_infreq(x)) %>% 
  add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_boxplot(fill=NA, lwd = 0.25) +
    geom_spoke(aes(x = as.numeric(x) + n/max(tmp_df$n)/2, radius = n/max(tmp_df$n), angle = pi, color = x),
               alpha = I(0.5), stat = "unique") +  # y = 0, radius = n for one-sided spoke plot
    coord_flip() +
    theme_minimal() +
    ggtitle(paste("distribution of", name[1], "grouped by", name[2], sep=" ")) 
fig <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

fig
```


## bivariate numeric score over numeric camis

- 

```{r}
# two variables, continuous x, continuous y, show trend and distribution
name = c('camis', 'score')
tmp_df <- df %>% rename(x = camis, y = score) %>% select(x, y) %>% add_count(x, y)

p1 <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_vline(aes(xintercept = quantile(tmp_df$x, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[2]), color = 'blue', alpha = 0.25) + 
    geom_hline(aes(yintercept = quantile(tmp_df$y, na.rm = TRUE)[4]), color = 'blue', alpha = 0.25) +
    geom_point(aes(size = n), alpha = 0.3, stat = "unique") + 
    geom_density2d(alpha = 0.5, bins = 4) +
    # geom_quantile(alpha = 0.4) +
    # geom_smooth(fill = "grey90") +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
p1 <- ggplotly(p1) %>% layout(xaxis = list(title = name[1]), yaxis = list(title = name[2]))

p2 <- tmp_df %>%
  ggplot(aes(x = x)) +
    stat_density(geom="line") + 
    theme_minimal()
p2 <- ggplotly(p2) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p3 <- tmp_df %>%
  ggplot(aes(x = 1, y = x)) +
    geom_boxplot() +
    theme_minimal() +
    coord_flip() 
p3 <- ggplotly(p3) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p4 <- tmp_df %>%
  ggplot(aes(x = y)) +
    stat_density(geom="line") + 
    coord_flip() +
    theme_minimal()
p4 <- ggplotly(p4) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

p5 <- tmp_df %>%
  ggplot(aes(x = 1, y = y)) +
    geom_boxplot() +
    theme_minimal() 
p5 <- ggplotly(p5) %>% layout(yaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''), 
                              xaxis = list(showticklabels = FALSE, showgrid = FALSE, title = ''))

# merge figures into one plot, via subplots, https://plotly-r.com/arranging-views.html
fig <- subplot(p2, plotly_empty(), plotly_empty(),
               p3, plotly_empty(), plotly_empty(),
               p1, p5, p4, 
               nrows = 3, margin = 0, heights = c(0.1, 0.035, 0.865), widths = c(0.92, 0.02, 0.06), 
               shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE) %>% layout()

fig
```


## bivarite categorical cuisine_description over categprocal boro

-

```{r}
name = c('cuisine_description', 'boro')
tmp_df <- df %>% rename(x = cuisine_description, y = boro) %>% select(x, y) 

fig <- tmp_df %>%
  ggplot(aes(x = x, y = y)) +
    geom_count() +
    theme_minimal() +
    ggtitle(paste("trend of", name[2], "over", name[1], sep=" "))
fig <- ggplotly(fig)

fig
```













